---
title: "Replication of 'Visual sense of number vs. sense of magnitude in humans and machines' by Testolin, A., Dolfi, S., Rochus, M. et al. (2020, Sci Rep)"
author: "Wenqing Cao (cwenqing@ucsd.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    toc: true
    toc_depth: 3
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

Link to github folder: https://github.com/psyc-201/testolin2020
Link to paradigm: https://psyc-201.github.io/testolin2020/testolin2020_experiment/experiment

## Introduction

Justification: 

It is widely believed that the approximate number system (ANS) supports formation of non-exact representation of quantities in adults, children, and even animals (Dehaene, 2011). ANS affords abilities like quickly discriminating two different numerical quantities or estimating the amount of items. The acuity of discrimintation is proportional to the ratio of the two quantities, and is found to predict future math learning performance. However, debate exists around whether numerosity is the main perceptual feature driving ANS, or numerosity is actually estimated from non-numerical visual features. This paper attempted to reconcile this debate by testing deep neural networks and human participants in the same task, adopting a stimulus space that could disentangle the contribution of numerical and non-numerical features. For this replication project, I will try to replicate the human experiment in this paper. It is related to my side research interest of probing ANS in vision language models. 

Methods: 

The research question is whether visual numerosity is a primary perceptual attribute or whether people estimate numerosity from continuous magnitudes. To test this, researchers adopted the dot array comparison task, where the participant sees two side-by-side dot arrays for a limited period of time, and then needs to indicate which array has a higher *number* of dots. To quantitatively estimate the contribution of non-numerical features, researchers constructed a 3-D orthogonal stimulus space which includes 3 dimensions: numerosity, size, and spacing. Specifically, they generated a database of 21970 images of dot arrays by picking 13 levels on each dimension, each level put evenly on a log scale. They then randomly sampled 300 pairs of dot images from the database as test stimuli, encompassing ratios from 0.5 to 0.9, biasing the harder ratios. Participants indicated which image in the pair had more dots by pressing left or right arrows on keyboard. 

They recruited volunteer college students (n = 40) as huamn subjects. Study session length for each participant ~ 30 minutes. The power of their human behavioral finding was strong, mean accuracy = 83%, with significant individual-level GLM fits (R² = 0.55, chi-square value = 191.14, p < 0.001). Coefficient fits for each dimension were significant for numerosity (t(39)=23.54, p<.001) and spacing (t(39)=7.21, p<.001), but not for size (t(39)=1.37, p=.18). 

The key analysis that I'm trying to replicate is the GLM fit on individual level, and the coefficient fit for each dimension. I will use the original stimuli (the 300 sampled pairs). I could recruit less participants (n = 25, 300 trials/30 mins each)   

What’s the measure of interest?
- Probability of choosing the dot array on the right side as having more dots 

What construct does it map to?
- Human adults’ perception of approximate numerosity

Is its estimate reported?
- effect size = 3.72, 1.14, 0.22 for numerosity, spacing, and size
- Unit: choose_right = 0/1 on a trial 

Any measure of variability?
- SE/error bars of beta coefficient


## Methods

### Power Analysis

Original effect size, power analysis for samples to achieve 80%, 90%, 95% power to detect that effect size.  Considerations of feasibility for selecting planned sample size.

A: The original study fitted a GLM model for each individaul, including 3 regressors numerosity, size, and spacing. The numerosity coefficient had the largest effect (t(39) = 23.54, p < 0.001); spacing coefficient also showed a significant effect (t(39) = 7.21, p < 0.001); size coefficient did not reveal a significant effect (t(39) = 1.37, p < 0.18). I calculated effect size by t/sqrt(n). Effect sizes for the numerosity, spacing, and sizing dimensions are 3.72 (very large), 1.14 (large), and 0.22 (small), respectively. Observed powers are 100% (numerosity), 100% (spacing), 24% (size), respectively. 

For numerosity dimension, 3, 4, 5 participants are needed to achieve 80%, 90%, 95% power. For spacing dimension, 8, 10, 13 participants are needed to achieve 80%, 90%, 95% power. For size dimension, 163, 219, 274 participants are needed to achieve 80%, 90%, 95% power. 

### Planned Sample

Planned sample size and/or termination rule, sampling frame, known demographics if any, preselection rules if any.

A: I plan to collect 30 samples, overpowered to account for a potentially high exclusion rate due to poor data quality or unfinished, given that the study is relatively long and tiring. Participants will be college undergraduate students in the local San Diego area, such that the sample's mean age and education level will roughly match the original study sample which consisted of 40 volunteer students (mean age = 23.7 years, range = 20-28, females = 80%). 

### Materials

All materials - can quote directly from original article - just put the text in quotations and note that this was followed precisely.  Or, quote directly and just point out exceptions to what was described in the original article.

A: "Images of size 200 × 200 pixels were generated by randomly placing white dots on a black background. For the discrimination task there were 13 levels of Numerosity (range 7–28), 13 levels of Size (range  2.6–10.4 pixels × 105) and 13 levels of Spacing (range 80–320 pixels × 105), evenly spaced on a logarithmic scale. For each selected point in the stimulus space 10 different images were generated by randomly varying dots displacement, resulting in a dataset of 21970 unique images. For the human experiment we randomly selected images from the dataset to create 300 image pairs with different magnitude ratios, oversampling the more difficult numerosity ratios (10% with ratio between 0.5 and 0.6; 20% with ratio between 0.6 and 0.7; 30% with ratio between 0.7 and 0.8; 40% with ratio between 0.8 and 0.9)." 

Because the authors did not publish code or data specifying the 300 image pairs used in the study, I will produce my own code to randomly select the 300 image pairs. The authors did publish the exact dimensions and point coordinates for all 21970 unique dot array images, so everything else will be equal.

### Procedure	

Can quote directly from original article - just put the text in quotations and note that this was followed precisely.  Or, quote directly and just point out exceptions to what was described in the original article.

A: "Stimuli were projected on a 19-inch color screen. Participants sat approximately 70 cm from the screen and placed their head on a chin rest. Participants were verbally instructed to select the stimulus with more dots, responding with the left and right arrows of the keyboard depending on its side of appearance (feedback was given only during few practice trials). The task consisted in 3 blocks of 100 trials each, for a total of 300 trials. Each trial began with a fixation cross at the center of the screen (500 ms), followed by the simultaneous presentation of two stimuli (250 ms), one at the right and one at the left of the cross with eccentricity of ~12 visual degrees, and then by two masks of black and white Gaussian noise in the same positions (150 ms). A black screen was then displayed until response, without time limit. After response, a pseudorandom inter-trial interval between 1250 and 1750 ms occurred."

### Analysis Plan

Can also quote directly, though it is less often spelled out effectively for an analysis strategy section.  The key is to report an analysis strategy that is as close to the original - data cleaning rules, data exclusion rules, covariates, etc. - as possible.  

**Clarify key analysis of interest here**  You can also pre-specify additional analyses you plan to do.

A: "A generalized linear model (with probit link function) was then fitted to the choice data of each  participant41, which was modeled as a function of the three regressors Numerosity, Size and Spacing" 

Following this procedure in the original paper, I will run a generalized linear model for each individual. The GLM model predicts the probability of choosing the image on the right from log ratio of numerosity, log ratio of spacing, and log ratio of size. The dependent measure is probability of choosing right image because the correct choice positions are randomized. Then I will run a t-test on all the coefficients of the individual GLM models. As exploratory analysis, I will fit a group-level general linear mixed effect model to all trials across participants to investigate the effect of numerosity, size, spacing, and random effect. The dependent measure will be the probability of choosing the image on the right. A GLM/GLMM is suitable because participants' responses were binary choices (left/right). 

### Differences from Original Study

Explicitly describe known differences in sample, setting, procedure, and analysis plan from original study.  The goal, of course, is to minimize those differences, but differences will inevitably occur.  Also, note whether such differences are anticipated to make a difference based on claims in the original article or subsequent published research on the conditions for obtaining the effect.

A: In my replication project, the sample size (n = 30) will be smaller than the previous study (n = 40). The setting and procedure will be the same. There will be a slight difference in stimuli as I will be re-sampling the 300 dot image pairs from their stimuli database. The images were upsized because the current study uses smaller screens (13-14 inch) compared to the original study (19 inch). 

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample

I recruited 27 participants on SONA for the formal experiment. After excluding participants whose overall accuracy is lower than 60%, I was left with 21 participants (mean age = 19.9, range = 18, 24) in the final data analysis. 

#### Differences from pre-data collection methods plan

Added exclusion criteria: I added exclusion criteria because when I looked at the data, I found out that some participants had extremely low accuracy rate or unreasonable response time, which are signs that they are not paying attention. Therefore participants are excluded from final analysis if their mean accuracy across all trials < 60%; specific trials are excluded if response time is outside of 50-2000 ms. 


## Results


### Data preparation

Data preparation following the analysis plan.

A: First, based on the exclusion criteria established, trials where a response is made below stimulus presentation time or where response time is over 2*SD from the participants' mean response time in equally difficult trials will be eliminated from the analysis. To achieve this, I will first merge all participants' data into a dataframe. Then I will apply a filter that only selects data rows where response time > stimulus presentation time (250 ms). I will also apply a filter that removes data where response time > mean(response time) + 2SD. I will make sure that columns include individual participant number, response time, 1/0 if correct on the trial, ratio in numerosity, ratio in size, ratio in spacing. 
	
```{r include=F}
### Data Preparation

#### Load Relevant Libraries and Functions
library(tidyverse)
library(broom)

#### Import data

select <- dplyr::select
filter <- dplyr::filter

metadata <- read_csv("/Users/annacao/Documents/github/testolin2020/testolin2020_experiment/paired_images/pairs_metadata.csv") %>%
    select(
      pair_num,
      pair_order,
      pair_ratio,
      left_image_id = image_id,
      left_numerosity = numerosity,
      left_TSA = TSA,
      left_FA = FA,
      left_CH = CH,
      left_cumArea = cumArea,
      left_role = role,   # anchor or partner
      position,   # which side this image is on
      greater_role,    # in this pair which role has more dots
      greater_side     # in this pair which side has more dots
    )

left_stim <- metadata %>%
    filter(position == "left") %>%
    select(-position)

right_stim <- metadata %>%
    filter(position == "right") %>%
    select(
      pair_order,
      right_image_id = left_image_id,
      right_numerosity = left_numerosity,
      right_TSA = left_TSA,
      right_FA = left_FA,
      right_CH = left_CH,
      right_cumArea = left_cumArea,
      right_role = left_role
    )

pairs_complete <- left_stim %>%
    left_join(right_stim, by = "pair_order") %>%
    select(
      pair_order,
      pair_ratio,
      left_image_id, left_numerosity, left_TSA, left_FA, left_CH, left_cumArea, left_role,
      right_image_id, right_numerosity, right_TSA, right_FA, right_CH, right_cumArea, right_role,
      greater_role, greater_side
    )

pairs_complete <- pairs_complete %>% 
  mutate(
    left_Size    = (left_TSA  / left_numerosity)^2,
    right_Size   = (right_TSA / right_numerosity)^2,
    left_Spacing = (left_FA   / left_numerosity)^2,
    right_Spacing= (right_FA  / right_numerosity)^2,
  )

pairs_complete <- pairs_complete %>% 
  mutate(
    log_num_ratio = log2(right_numerosity) - log2(left_numerosity),
    log_size_ratio = log2(right_Size) - log2(left_Size),
    log_spacing_ratio = log2(right_Spacing) - log2(left_Spacing)
  )

pairs_complete <- pairs_complete %>% 
  mutate(
    congruent_spacing = sign(log_num_ratio) == sign(log_spacing_ratio),
    congruent_size    = sign(log_num_ratio) == sign(log_size_ratio)
  )


# load participant data
load_participant_data <- function(participant_path) {
  participant_data <- read_csv(participant_path,
      col_types = cols(
      participant_id = col_character(),  
      screen_size_category = col_character(),
      .default = col_guess()
      ),
      show_col_types = FALSE
      )%>%
    filter(task == "numerosity_discrimination") %>%
    select(
      pair_order,
      pair_num,
      pair_ratio,
      rt,
      response,
      trial_index,
      time_elapsed,
      participant_id,
      participant_age,
      block,
      any_of(c(
        "screen_size_category",
        "screen_width_pixels",
        "screen_height_pixels",
        "window_width_pixels",
        "window_height_pixels",
        "pixel_ratio"
      ))
    )
  return(participant_data)
}

data_path <- list.files(
  path = "/Users/annacao/Documents/github/testolin2020/data", 
  pattern = "*.csv", 
  full.names = TRUE
)

all_data <- map_dfr(data_path, function(file) {
  data <- load_participant_data(file)
  name <- unique(data$participant_id)
  message("Loaded participant: ", name)
  data
})
  

#### Data exclusion / filtering

#### Prepare data for analysis - create columns etc.

# initial sample size
n_total <- n_distinct(all_data$participant_id)
n_trials_total <- nrow(all_data)

cat("INITIAL SAMPLE:\n")
cat("  Total participants:", n_total, "\n")
cat("  Total trials:", n_trials_total, "\n\n")

# Exclude participants who didn't complete the experiment
expected_trials <- 300

participant_trial_counts <- all_data %>%
  group_by(participant_id) %>%
  summarise(
    n_trials = n(),
    .groups = "drop"
  )

incomplete_participants <- participant_trial_counts %>%
  filter(n_trials < expected_trials) %>%
  pull(participant_id)

if (length(incomplete_participants) > 0) {
  cat("STEP 1: EXCLUDE INCOMPLETE PARTICIPANTS\n")
  cat("  Participants with < ", expected_trials, " trials: ", 
      length(incomplete_participants), "\n", sep = "")
  
  cat("  Excluded IDs:\n")
  for (pid in incomplete_participants) {
    n_trials <- participant_trial_counts$n_trials[participant_trial_counts$participant_id == pid]
    cat("    -", pid, "(", n_trials, "trials )\n")
  }
  
  all_data <- all_data %>%
    filter(!participant_id %in% incomplete_participants)
  
  cat("  Remaining participants:", n_distinct(all_data$participant_id), "\n\n")
} else {
  cat("STEP 1: EXCLUDE INCOMPLETE PARTICIPANTS\n")
  cat("  All participants completed the experiment\n")
  cat("  No exclusions\n\n")
}

# participant correctness column 
compute_correctness <- function(merged_data) {
  merged_data %>%
    mutate(
      correct_response = case_when(
        greater_side == "left" ~ "f",
        greater_side == "right" ~ "j",
        TRUE ~ NA_character_
      ),
      correct = (response == correct_response)
    )
}

  merged <- all_data %>%
    left_join(pairs_complete, by = c("pair_order")) %>%
    compute_correctness() %>%
    
    group_by(participant_id, block) %>%
    mutate(trial_in_block = row_number()) %>%
    ungroup()
  
  qc <- merged %>%
    group_by(participant_id) %>%
    summarise(
      n_trials   = n(),
      n_missing  = sum(is.na(left_numerosity)),
      mean_correct = mean(correct, na.rm = TRUE),
      .groups = "drop"
    )
  
  qc %>% 
    rowwise() %>%
    mutate(
      msg = sprintf(
        "Participant %s: %d trials, %d missing merges, %.1f%% correct",
        participant_id, n_trials, n_missing, 100 * mean_correct
      )
    ) %>%
    pull(msg) %>%
    walk(message)

# Exclude trials with unrealistic RTs
rt_min <- 50   # minimum RT in ms
rt_max <- 2000  # maximum RT in ms

cat("  RT cutoffs: ", rt_min, "ms - ", rt_max, "ms\n", sep = "")

n_trials_before_rt <- nrow(merged)

merged_rt_filtered <- merged %>%
  filter(rt >= rt_min & rt <= rt_max)

n_trials_excluded_rt <- n_trials_before_rt - nrow(merged_rt_filtered)
percent_excluded_rt <- round(n_trials_excluded_rt / n_trials_before_rt * 100, 2)

cat("  Trials excluded for RT:", n_trials_excluded_rt, 
    "(", percent_excluded_rt, "%)\n", sep = "")
cat("  Remaining trials:", nrow(merged_rt_filtered), "\n\n")

# Exclude participants with accuracy below threshold (e.g., 60%)
participant_accuracy <- merged_rt_filtered %>%
  group_by(participant_id) %>%
  summarise(
    mean_accuracy = mean(correct, na.rm = TRUE),
    n_trials = n(),
    .groups = "drop"
  )

accuracy_threshold <- 0.60  # 60% accuracy

cat("  Accuracy threshold:", accuracy_threshold * 100, "%\n", sep = "")

low_accuracy_participants <- participant_accuracy %>%
  filter(mean_accuracy < accuracy_threshold)

if (nrow(low_accuracy_participants) > 0) {
  cat("  Participants with accuracy <", accuracy_threshold * 100, "%: ", 
      nrow(low_accuracy_participants), "\n", sep = "")
  
  cat("  Excluded IDs:\n")
  low_accuracy_participants %>%
    rowwise() %>%
    mutate(
      msg = sprintf("    - %s (%.1f%% accuracy, %d trials)",
                   participant_id, mean_accuracy * 100, n_trials)
    ) %>%
    pull(msg) %>%
    walk(cat, "\n")
  
  low_accuracy_ids <- low_accuracy_participants$participant_id
  
  merged_final <- merged_rt_filtered %>%
    filter(!participant_id %in% low_accuracy_ids)
  
  cat("  Remaining participants:", n_distinct(merged_final$participant_id), "\n\n")
} else {
  cat("  All participants meet accuracy threshold\n")
  cat("  No exclusions\n\n")
  merged_final <- merged_rt_filtered
}

cat("==========================================================\n")
cat("FINAL SAMPLE SUMMARY\n")
cat("==========================================================\n\n")

n_participants_final <- n_distinct(merged_final$participant_id)
n_trials_final <- nrow(merged_final)

n_participants_excluded <- n_total - n_participants_final
n_trials_excluded <- n_trials_total - n_trials_final

cat("PARTICIPANTS:\n")
cat("  Initial:", n_total, "\n")
cat("  Excluded:", n_participants_excluded, "\n")
cat("  Final:", n_participants_final, "\n\n")

cat("TRIALS:\n")
cat("  Initial:", n_trials_total, "\n")
cat("  Excluded:", n_trials_excluded, 
    "(", round(n_trials_excluded / n_trials_total * 100, 2), "%)\n", sep = "")
cat("  Final:", n_trials_final, "\n\n")

# Quality check on final data
qc_final <- merged_final %>%
  group_by(participant_id) %>%
  summarise(
    n_trials = n(),
    mean_accuracy = mean(correct, na.rm = TRUE),
    mean_rt = mean(rt, na.rm = TRUE),
    median_rt = median(rt, na.rm = TRUE),
    .groups = "drop"
  )

cat("FINAL DATA QUALITY:\n")
cat("  Mean accuracy:", round(mean(qc_final$mean_accuracy) * 100, 1), "%\n", sep = "")
cat("  Mean RT:", round(mean(qc_final$mean_rt), 0), "ms\n", sep = "")
cat("  Median RT:", round(median(qc_final$median_rt), 0), "ms\n\n", sep = "")

cat("==========================================================\n\n")

merged <- merged_final

mean_age_merge <- mean(merged$participant_age)
range_age_merge <- range(merged$participant_age)

```

### Confirmatory analysis

I will fit the GLM model to each individual. Specifically, fixed effects will be log-ratios of numerosity, size, spacing. From the resulting model I will then run t-test to compare the effect sizes of numerosity, size, spacing, to determine which/whether dimension is the main predictor of performance (or which dimension dominates approximate number discrimination).  

```{r}

summary_stats <- merged %>%
    group_by(participant_id) %>%
    summarise(
      n_trials = n(),
      accuracy = mean(correct, na.rm = TRUE),
      mean_rt = mean(rt, na.rm = TRUE),
      median_rt = median(rt, na.rm = TRUE),
      .groups = "drop"
    )

summary_stats

#### confirmatory GLM
 merged <- merged %>%
  mutate(
    # 1 = chose right (J), 0 = chose left (F)
    chose_right = if_else(response == "j", 1L, 0L)
  )

fit_glm_per_participant <- function(data) {
  data %>%
    group_by(participant_id) %>%
    group_modify(~{
      model <- glm(
        chose_right ~ log_num_ratio + log_size_ratio + log_spacing_ratio,
        data = .x,
        family = binomial(link = "probit")
      )
      tidy(model) %>%
        mutate(
          participant_id = unique(.x$participant_id),
          n_trials = nrow(.x),
          pseudo_r2 = 1 - model$deviance / model$null.deviance
        )
    }) %>%
    ungroup()
}

glm_results <- fit_glm_per_participant(merged)

glm_results


#### one-sample t-test on each dimension beta

# Test betaNum
t_test_num <- t.test(beta_wide$betaNum, mu = 0)
cat("βNum (Numerosity):\n")
cat("  Mean:", round(mean(beta_wide$betaNum), 3), "\n")
cat("  SD:", round(sd(beta_wide$betaNum), 3), "\n")
cat("  t(", t_test_num$parameter, ") = ", round(t_test_num$statistic, 3), 
    ", p ", ifelse(t_test_num$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_num$p.value, 3))), "\n",
    sep = "")
cat("  95% CI: [", round(t_test_num$conf.int[1], 3), ", ", 
    round(t_test_num$conf.int[2], 3), "]\n\n", sep = "")

# Test betaSize
t_test_size <- t.test(beta_wide$betaSize, mu = 0)
cat("βSize (Size):\n")
cat("  Mean:", round(mean(beta_wide$betaSize), 3), "\n")
cat("  SD:", round(sd(beta_wide$betaSize), 3), "\n")
cat("  t(", t_test_size$parameter, ") = ", round(t_test_size$statistic, 3), 
    ", p ", ifelse(t_test_size$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_size$p.value, 3))), "\n",
    sep = "")
cat("  95% CI: [", round(t_test_size$conf.int[1], 3), ", ", 
    round(t_test_size$conf.int[2], 3), "]\n\n", sep = "")

# Test betaSpacing
t_test_spacing <- t.test(beta_wide$betaSpacing, mu = 0)
cat("βSpacing (Spacing):\n")
cat("  Mean:", round(mean(beta_wide$betaSpacing), 3), "\n")
cat("  SD:", round(sd(beta_wide$betaSpacing), 3), "\n")
cat("  t(", t_test_spacing$parameter, ") = ", round(t_test_spacing$statistic, 3), 
    ", p ", ifelse(t_test_spacing$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_spacing$p.value, 3))), "\n",
    sep = "")
cat("  95% CI: [", round(t_test_spacing$conf.int[1], 3), ", ", 
    round(t_test_spacing$conf.int[2], 3), "]\n\n", sep = "")

```


*Side-by-side graph with original graph is ideal here*

```{r}
beta_wide <- glm_results %>%
  filter(term %in% c("log_num_ratio", "log_size_ratio", "log_spacing_ratio")) %>%
  select(participant_id, term, estimate) %>%
  pivot_wider(
    names_from  = term,
    values_from = estimate
  ) %>%
  rename(
    betaNum     = log_num_ratio,
    betaSize    = log_size_ratio,
    betaSpacing = log_spacing_ratio
  )

beta_long <- beta_wide %>%
  pivot_longer(cols = c(betaNum, betaSize, betaSpacing),
               names_to = "dimension",
               values_to = "beta")


#### scatter plots for beta_numerosity, beta_spacing, beta_size
scatter_theme <- theme_classic() +
  theme(
    axis.title = element_text(size = 12),
    axis.text  = element_text(size = 10)
  )

# Panel 1: βNum vs βSize
p_num_size <- ggplot(beta_wide, aes(x = betaSize, y = betaNum)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_vline(xintercept = 0, colour = "grey70") +
  geom_abline(intercept = 0, slope = 1, colour = "grey85") +
  geom_abline(intercept = 0, slope = -1, colour = "grey85") +
  geom_point(shape = 21, colour = "red", fill = NA, size = 2) +
  coord_equal(xlim = c(-6, 6), ylim = c(-6, 6)) +
  labs(
    x = expression(beta[Size]),
    y = expression(beta[Num])
  ) +
  scatter_theme

p_num_size

# Panel 2: βNum vs βSpacing
p_num_spacing <- ggplot(beta_wide, aes(x = betaSpacing, y = betaNum)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_vline(xintercept = 0, colour = "grey70") +
  geom_abline(intercept = 0, slope = 1, colour = "grey85") +
  geom_abline(intercept = 0, slope = -1, colour = "grey85") +
  geom_point(shape = 21, colour = "red", fill = NA, size = 2) +
  coord_equal(xlim = c(-6, 6), ylim = c(-6, 6)) +
  labs(
    x = expression(beta[Spacing]),
    y = expression(beta[Num])
  ) +
  scatter_theme

p_num_spacing

# Panel 3: βSpacing vs βSize
p_spacing_size <- ggplot(beta_wide, aes(x = betaSize, y = betaSpacing)) +
  geom_hline(yintercept = 0, colour = "grey70") +
  geom_vline(xintercept = 0, colour = "grey70") +
  geom_abline(intercept = 0, slope = 1, colour = "grey85") +
  geom_abline(intercept = 0, slope = -1, colour = "grey85") +
  geom_point(shape = 21, colour = "red", fill = NA, size = 2) +
  coord_equal(xlim = c(-6, 6), ylim = c(-6, 6)) +
  labs(
    x = expression(beta[Size]),
    y = expression(beta[Spacing])
  ) +
  scatter_theme

p_spacing_size

#### individual GLM psychometric curve
library(dplyr)
library(ggplot2)
library(broom)

# pick extreme threshold
merged %>%
  select(log_size_ratio, log_spacing_ratio) %>%
  pivot_longer(everything(), names_to = "ratio_type", values_to = "value") %>%
  ggplot(aes(x = value, fill = ratio_type)) +
  geom_histogram(bins = 50, alpha = 0.6, position = "identity") +
  geom_vline(xintercept = c(-0.7, 0.7), linetype = "dashed", color = "red") +
  labs(title = "Distribution of Size and Spacing Ratios",
       x = "Log2 Ratio",
       y = "Count") +
  theme_classic()

# Check what percentile 0.7 represents
quantile(abs(merged$log_size_ratio), probs = c(0.5, 0.67, 0.75, 0.8, 0.9))
quantile(abs(merged$log_spacing_ratio), probs = c(0.5, 0.67, 0.75, 0.8, 0.9))

size_threshold <- quantile(abs(merged$log_size_ratio), probs = 0.75)
spacing_threshold <- quantile(abs(merged$log_spacing_ratio), probs = 0.75)

cat("Size threshold (75th percentile):", size_threshold, "\n")
cat("Spacing threshold (75th percentile):", spacing_threshold, "\n")


# plot individual probit graph
plot_psychometric_participant <- function(df, pid, extreme_thresh = 0.15) {
  
   d <- df %>%
    filter(participant_id == pid) %>%
    mutate(
      # Flag extreme trials
      size_extreme = abs(log_size_ratio) >= extreme_thresh,
      spacing_extreme = abs(log_spacing_ratio) >= extreme_thresh
    )
  
  model_full <- glm(
    chose_right ~ log_num_ratio + log_size_ratio + log_spacing_ratio,
    data = d,
    family = binomial(link = "probit")
  )
  
  cat("\n=== Model Coefficients ===\n")
  print(summary(model_full)$coefficients)
  
  size_cong_data <- d %>% filter(size_extreme, congruent_size)
  size_incong_data <- d %>% filter(size_extreme, !congruent_size)
  spacing_cong_data <- d %>% filter(spacing_extreme, congruent_spacing)
  spacing_incong_data <- d %>% filter(spacing_extreme, !congruent_spacing)

  typical_size_extreme <- median(abs(d$log_size_ratio[d$size_extreme]))
  typical_spacing_extreme <- median(abs(d$log_spacing_ratio[d$spacing_extreme]))
  
  cat("\n=== Typical Extreme Values ===\n")
  cat("Size extreme:", typical_size_extreme, "\n")
  cat("Spacing extreme:", typical_spacing_extreme, "\n")
  
  x_grid <- seq(-1, 1, length.out = 300)
  
  pred_all <- data.frame(
    log_num_ratio = x_grid,
    log_size_ratio = 0,
    log_spacing_ratio = 0
  ) %>%
    mutate(
      p_hat = predict(model_full, newdata = ., type = "response"),
      condition = "All trials",
      color_group = "black",
      linetype_group = "solid"
    )
  
  pred_size_cong <- data.frame(
    log_num_ratio = x_grid,
    log_size_ratio = sign(x_grid) * typical_size_extreme,
    log_spacing_ratio = 0
  ) %>%
    mutate(
      p_hat = predict(model_full, newdata = ., type = "response"),
      condition = "Size congruent",
      color_group = "red",
      linetype_group = "dashed"
    )
  
  # RED DOTTED: Size incongruent
  # When num_ratio > 0: size_ratio < 0 (incongruent)
  # When num_ratio < 0: size_ratio > 0 (incongruent)
  # So: size_ratio = -sign(num_ratio) × typical_size_extreme
  pred_size_incong <- data.frame(
    log_num_ratio = x_grid,
    log_size_ratio = -sign(x_grid) * typical_size_extreme,
    log_spacing_ratio = 0
  ) %>%
    mutate(
      p_hat = predict(model_full, newdata = ., type = "response"),
      condition = "Size incongruent",
      color_group = "red",
      linetype_group = "dotted"
    )
  
  # GREEN DASHED: Spacing congruent
  pred_spacing_cong <- data.frame(
    log_num_ratio = x_grid,
    log_size_ratio = 0,
    log_spacing_ratio = sign(x_grid) * typical_spacing_extreme
  ) %>%
    mutate(
      p_hat = predict(model_full, newdata = ., type = "response"),
      condition = "Spacing congruent",
      color_group = "green4",
      linetype_group = "dashed"
    )
  
  # GREEN DOTTED: Spacing incongruent
  pred_spacing_incong <- data.frame(
    log_num_ratio = x_grid,
    log_size_ratio = 0,
    log_spacing_ratio = -sign(x_grid) * typical_spacing_extreme
  ) %>%
    mutate(
      p_hat = predict(model_full, newdata = ., type = "response"),
      condition = "Spacing incongruent",
      color_group = "green4",
      linetype_group = "dotted"
    )
  
  # Combine all predictions
  all_preds <- bind_rows(
    pred_all,
    pred_size_cong,
    pred_size_incong,
    pred_spacing_cong,
    pred_spacing_incong
  )
  
  bin_width <- 0.1
  
  create_binned_points <- function(data, cond_name, col) {
    if (nrow(data) == 0) return(NULL)
    
    data %>%
      mutate(num_bin = round(log_num_ratio / bin_width) * bin_width) %>%
      group_by(num_bin) %>%
      summarise(
        p_right = mean(chose_right),
        n = n(),
        .groups = "drop"
      ) %>%
      filter(n >= 2) %>%  # Only show bins with at least 2 trials
      mutate(
        condition = cond_name,
        color_group = col
      )
  }
  
  # All trials
  emp_all <- create_binned_points(d, "All trials", "black")
  
  # Size conditions
  emp_size_cong <- create_binned_points(size_cong_data, "Size congruent", "red")
  emp_size_incong <- create_binned_points(size_incong_data, "Size incongruent", "red")
  
  # Spacing conditions
  emp_spacing_cong <- create_binned_points(spacing_cong_data, "Spacing congruent", "green4")
  emp_spacing_incong <- create_binned_points(spacing_incong_data, "Spacing incongruent", "green4")
  
  # Combine all empirical points
  all_points <- bind_rows(
    emp_all,
    emp_size_cong,
    emp_size_incong,
    emp_spacing_cong,
    emp_spacing_incong
  )
  
  p <- ggplot() +
    # Plot lines first (behind points)
    geom_line(
      data = all_preds,
      aes(x = log_num_ratio, y = p_hat, 
          color = color_group, linetype = linetype_group,
          group = interaction(color_group, linetype_group)),
      linewidth = 1
    ) +
    
    # Plot points on top
    geom_point(
      data = all_points,
      aes(x = num_bin, y = p_right, 
          color = color_group,
          shape = condition),
      size = 3,
      stroke = 1.2
    ) +
    
    # Manual scales
    scale_color_identity() +
    scale_linetype_identity() +
    
    scale_shape_manual(
      name = "",
      values = c(
        "All trials" = 1,              # open circle
        "Size congruent" = 1,          # open circle
        "Size incongruent" = 1,        # open circle
        "Spacing congruent" = 1,       # open circle
        "Spacing incongruent" = 1      # open circle
      )
    ) +
    
    coord_cartesian(xlim = c(-1, 1), ylim = c(0, 1)) +
    
    labs(
      x = expression("Numerosity Ratio (log"[2]*")"),
      y = "p(Choose Right)",
      title = paste("Participant", pid)
    ) +
    
    theme_classic() +
    theme(
      legend.position = "right",
      legend.text = element_text(size = 10),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      plot.title = element_text(size = 13, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )
  
  cat("\n=== Sample Sizes ===\n")
  cat("All trials:", nrow(d), "\n")
  cat("Size congruent:", nrow(size_cong_data), "\n")
  cat("Size incongruent:", nrow(size_incong_data), "\n")
  cat("Spacing congruent:", nrow(spacing_cong_data), "\n")
  cat("Spacing incongruent:", nrow(spacing_incong_data), "\n")
  
  cat("\n=== Binned Points ===\n")
  print(table(all_points$condition))
  
  return(p)
}

example_id <- unique(merged$participant_id)[3]

plot_psychometric_participant(merged, example_id)

plot_psychometric_participant(merged, unique(merged$participant_id)[1])

plot_psychometric_participant(merged, unique(merged$participant_id)[2])

plot_psychometric_participant(merged, unique(merged$participant_id)[3])

plot_psychometric_participant(merged, unique(merged$participant_id)[4])

plot_psychometric_participant(merged, unique(merged$participant_id)[5])

```


### Exploratory analyses

The result shows that both numerosity and spacing had a large effect, whereas size had a smaller effect. I will run a GLMM at group level, and also run paired t-tests for the numerosity, spacing, and size effects. 

```{r}
#### GLMM analysis
#install.packages(c("lme4", "broom.mixed", "performance"))
library(lme4)
library(broom.mixed)
library(dplyr)

glmm_model <- glmer(
  chose_right ~ log_num_ratio + log_size_ratio + log_spacing_ratio + (1 | participant_id),
  data = merged,
  family = binomial(link = "probit"),
  control = glmerControl(optimizer = "bobyqa")
)

summary(glmm_model)

glmm_results <- broom.mixed::tidy(glmm_model, effects = "fixed")

glmm_results



#### paired t-test
# βNum vs βSize
t_test_num_vs_size <- t.test(beta_wide$betaNum, beta_wide$betaSize, paired = TRUE)
cat("βNum vs βSize:\n")
cat("  Mean difference:", round(mean(beta_wide$betaNum - beta_wide$betaSize), 3), "\n")
cat("  t(", t_test_num_vs_size$parameter, ") = ", round(t_test_num_vs_size$statistic, 3), 
    ", p ", ifelse(t_test_num_vs_size$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_num_vs_size$p.value, 3))), "\n",
    sep = "")
cat("  Interpretation:", ifelse(t_test_num_vs_size$p.value < 0.05,
                                "Numerosity effect is significantly larger than Size effect",
                                "No significant difference"), "\n\n")

# βNum vs βSpacing
t_test_num_vs_spacing <- t.test(beta_wide$betaNum, beta_wide$betaSpacing, paired = TRUE)
cat("βNum vs βSpacing:\n")
cat("  Mean difference:", round(mean(beta_wide$betaNum - beta_wide$betaSpacing), 3), "\n")
cat("  t(", t_test_num_vs_spacing$parameter, ") = ", round(t_test_num_vs_spacing$statistic, 3), 
    ", p ", ifelse(t_test_num_vs_spacing$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_num_vs_spacing$p.value, 3))), "\n",
    sep = "")
cat("  Interpretation:", ifelse(t_test_num_vs_spacing$p.value < 0.05,
                                "Numerosity effect is significantly larger than Spacing effect",
                                "No significant difference"), "\n\n")

# βSize vs βSpacing
t_test_size_vs_spacing <- t.test(beta_wide$betaSize, beta_wide$betaSpacing, paired = TRUE)
cat("βSize vs βSpacing:\n")
cat("  Mean difference:", round(mean(beta_wide$betaSize - beta_wide$betaSpacing), 3), "\n")
cat("  t(", t_test_size_vs_spacing$parameter, ") = ", round(t_test_size_vs_spacing$statistic, 3), 
    ", p ", ifelse(t_test_size_vs_spacing$p.value < 0.001, "< 0.001", 
                   paste("=", round(t_test_size_vs_spacing$p.value, 3))), "\n",
    sep = "")
cat("  Interpretation:", ifelse(t_test_size_vs_spacing$p.value < 0.05,
                                "Size effect is significantly larger than Spacing effect",
                                "No significant difference"), "\n\n")

cat("========== SUMMARY TABLE ==========\n\n")

summary_table <- data.frame(
  Dimension = c("Numerosity", "Size", "Spacing"),
  Mean = c(mean(beta_wide$betaNum), mean(beta_wide$betaSize), mean(beta_wide$betaSpacing)),
  SD = c(sd(beta_wide$betaNum), sd(beta_wide$betaSize), sd(beta_wide$betaSpacing)),
  SE = c(sd(beta_wide$betaNum)/sqrt(nrow(beta_wide)), 
         sd(beta_wide$betaSize)/sqrt(nrow(beta_wide)),
         sd(beta_wide$betaSpacing)/sqrt(nrow(beta_wide))),
  t_statistic = c(t_test_num$statistic, t_test_size$statistic, t_test_spacing$statistic),
  df = c(t_test_num$parameter, t_test_size$parameter, t_test_spacing$parameter),
  p_value = c(t_test_num$p.value, t_test_size$p.value, t_test_spacing$p.value),
  significant = c(t_test_num$p.value < 0.05, t_test_size$p.value < 0.05, t_test_spacing$p.value < 0.05)
)

print(summary_table, digits = 4)

```


## Discussion

### Summary of Replication Attempt

This replication successfully confirmed the primary finding of Testolin et al. (2020): human adults use both numerical and non-numerical visual features when making rapid numerosity judgments. Our confirmatory analysis using individual-level GLM fits revealed significant effects of numerosity (mean β = 2.815, t(20) = 12.449, p < 0.001) and spacing (mean β = 0.18, t(20) = 4.141, p < 0.001), but not size (mean β = 0.109, t(20) = 2.262,, p = 0.035). These results closely parallel the original study's findings (numerosity: t(39) = 23.54; spacing: t(39) = 7.21; size: t(39) = 1.37), with nearly identical effect sizes and the same pattern of statistical significance. I therefore conclude that this is a successful replication of the original finding. The effect sizes are large and still well-powered even with half of the original sample size. Results support that numerosity is the primary dimension of estimation in visual numerosity perception, with spacing also playing a role on the performance. 

In exploratory analysis, the group-level GLMM analysis corroborated these individual-level results, showing robust main effects of numerosity (β = 2.499, z = 36.97, p < 0.001) and spacing (β = 0.123 z = 5.32, p < 0.001), but also significant effect of size (β = 0.145, z = 5.94, p < 0.001). It is interesting how size is not significant when running a t-test of all size coefficient of the individual GLMs, but becomes significant in the group-level GLMM. Paired t-tests, on the other hand, shows that when comparing the individual GLM coefficients, numerosity coefficients are significantly different from the spacing coefficient and the size coefficient, but spacing coefficient and the size coefficient are not significantly different. 

### Commentary

These findings contribute to the ongoing debate about the nature of the approximate number system. Our replication confirms that while humans can extract numerical information from visual arrays rapidly and accurately, this process is not purely "number-based" but rather involves integration of multiple visual features. The significant spacing effect—replicated across both studies suggests that spatial distribution of objects influences numerosity judgments even when participants are explicitly instructed to judge number. This is consistent with theories proposing that the ANS relies on an ensemble coding mechanism that integrates multiple visual statistics rather than extracting a pure numerosity signal. However, the dominance of the numerosity coefficient (roughly 3-4 times larger than the spacing coefficient) indicates that while non-numerical features play a role, number remains the primary perceptual feature driving these judgments.

As a next step, it would be valuable to test whether computational models trained on similar stimuli develop the same feature dependencies observed in humans, as this could provide insights into the mechanisms underlying visual numerosity perception.




